---
title: "Week 4 Part 1 - CIs & t-tests"
author: "Larissa Neilson"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 1. Confidence intervals

```{r}
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)

t.test(otter_length)
```
Interpretation: "The mean length of sea otters (n = x) is 42.22 inches with a SD of x. 95% interval is 34.7 - 49.8 inches"

## T-test (1-sample  t-test)

POV - You see the claim: "mean otter length is 50 inches".
I will use my sample to test this claim!

```{r}
otter_test <- t.test(x = otter_length, mu = 50)
otter_test
```
This tells us: "If the population mean really is 50 inches, then there is a 4.44% chance that I could have randomly taken a sample from that population that has a mean that is at least as different from 50 inches as my sample mean is in either direction."

Do not think of the p-value as significant or not!
Interesting thing to consider is how different the mean is from the claim of 50 inches.

Ask: Is it possible that my sample was drawn from a population that has a mean of 50 inches? Yes it's possible, but not probable (4.44% chance).

Are differences in the two sample means so different that I actually think they were drawn from populations with different means?

## Two-sample t-test

```{r}
desert_bighorns <- c(32, 44, 18, 26, 50, 33, 42, 20)
sierra_bighorns <- c(28, 31, 40, 42, 26, 29, 31)

t.test(x = desert_bighorns, y = sierra_bighorns)
```
This test does not assume equal variance ^

CI can be negative because y-x, x-y
We expect the difference in mean horn length between species to be within the CI range for 95% of the samples that we took <--- this info is less meaningful in this case!

P-value is a probability - It tells me that if these samples are drawn from populations with the same mean horn length (null hypothesis), then there is almost a 90% chance that I could have taken two random samples from those populations and found samples at least as different as the samples that I have here.

This tells me that there is a really GOOD chance that I could have retrieved my samples from populations with the same mean.

Conclusion: I do not think I have enough evidence to reject the null hypothesis.

## END PART 1

## Part 2: Two-sample t-test

```{r}
compact_suv <- mpg %>%
  filter(class %in% c("compact", "suv"))
```

## Exploratory data visualization

Exploratory histograms:
```{r}
ggplot(data = compact_suv, aes(x = cty)) +
  geom_histogram(bins = 15) +
  facet_wrap(~class)
```
No major concerns with histograms!

## Quantile-quantile plot

```{r}
ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~class)
```
No major concerns about distribution overall. Possible positive skew in compact cars.

## Descriptive statistics

```{r}
car_stats <- compact_suv %>%
  group_by(class) %>% 
  summarize(mean_city = mean(cty),
            sd_city = sd(cty),
            n = n())
car_stats
```
## 2-sample t-test

Create vectors containing observations for each sample (city gas mileage)

```{r}
compact_sample <- compact_suv %>%
  filter(class == "compact") %>%
  pull(cty)

suv_sample <- compact_suv %>%
  filter(class == "suv") %>%
  pull(cty)

t.test(x = compact_sample, suv_sample)
```

Null: drawn from populations with same exact mean

Close to 0% chance of randomly drawing means...

Conclusion: samples *WERE* drawn from populations with different means.